---
title: |
  | STATS 3DS3
  | Homework assignment 1
  | Data visualization (Lecture 1, 2, 3, and 4)
author: 
  - "Ling Lu (with Weijiang Song's and Lan Wang's help)"
date: "01/21/2021"
output: 
  pdf_document:
    includes:
      in_header: header.tex
bibliography: STATS3DS3.bib
---

**You may discuss homework problems with other students, but you have to prepare the written assignments yourself.**

**Please combine all your answers, the computer code and the figures into one PDF file, and submit a copy to Crowdmark.**

Please use **newpage** to write solution for each part of a question.

Please choose the pages for each part of a question in Crowdmark. 

**Grading scheme: $\left\lbrace 0, 1, 2\right\rbrace$ points per question, total of 44.** We will convert this to 100\%.

Homework assignment 1 due date: \rc Friday, January 29th, 2021 at 10:00 PM.\bc


\newpage


##  Question 1 (R basics) 

**RDS** 4.4 exercises 

1. Why the following code doesn't work? Correct it and run it. Show the R commands and print the output. 

```{r }
my_variable <- 10
my_variable
```




\newpage

2. We can use `class()` function to find the type of object. Find the type of `mpg` and `diamonds` data in ggplot2 package. Show the R commands. When we load the package `tidyverse`, the output shows warnings and messages. We can remove these warnings and messages in the output by setting chunk options as {r message = FALSE, warnings = FALSE}.

```{r message=FALSE, warnings=FALSE}
library(tidyverse)
library(ggplot2)
class(mpg)
class(diamonds)
```

\newpage

3. `mpg` and `diamonds` data are in ggplot2 package. Tweak each of the following R commands so that they run correctly. In the output, hide the R commands. We can use chunk option `echo=FALSE` to hide R commands. In the output, center the plot using chunk option `fig.align = 'center'`.
   
```{r echo=FALSE, fig.align='center'}
library(tidyverse)

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))

filter(mpg, cyl == 8)
filter(diamonds, carat > 3)
```
  


\newpage

## Question 2 (Data set exploration)

`Titanic {datasets}` data set provides information on the fate of passengers on the fatal maiden voyage of the ocean liner ‘Titanic’, summarized according to economic status (class), sex, age and survival.


1) Assign `Titanic` data set from `datasets` package to `titanic`. Show the R commands.

```{r}
library(datasets)
titanic <-datasets::Titanic
```


\newpage

2) Find out the type of `titanic`. In the output, hide the R commands.
```{r echo=FALSE}
class(titanic)
```

\newpage

3) `titanic` is not a data frame. Convert `titanic` to a tibble using `as_tibble()`. Show the R commands. 
```{r}
titanic1 <-as_tibble(titanic)
```

\newpage

4) Print the first 10 observations from the `titanic` tibble. In the output, hide the R commands. We see a variable called `n`. This will be the frequency of each group. For example, the first row is read as no male child in the 1st class survived. 

```{r echo=FALSE}
head(titanic1,10)
```


\newpage

5) Identify the type of each variable in the Titanic data. Show the R commands. 

```{r }
sapply(titanic1, class)
```

\newpage


## Question 3 (Data transformation and visualization)

We will use [\blc Spotify Tracks DB dataset from Kaggle\bc](https://www.kaggle.com/zaheenhamidani/ultimate-spotify-tracks-db), which was curated using the Spotify Web API. The Web API provides a variety of information about each track, using the Spotify URI and the Spotify ID as track identifiers.

1) Visit the [\blc website\bc](https://www.kaggle.com/zaheenhamidani/ultimate-spotify-tracks-db) and download the file `SpotifyFeatures.csv`. Read data into R using `readr::read_csv()` and call the tibble as `spotify`. When we use `readr::read_csv()`, we would specify that `col_names = TRUE`.

```{r message = FALSE}
library(readr)
getwd()
spotify <- readr::read_csv("SpotifyFeatures.csv", col_names = TRUE)
```


\newpage

2) How many observations and variables are in the data set? Use an inline R command.

```{r}
observations <- nrow(spotify)
variables <-ncol(spotify)
```
there are `r observations` observations and `r variables` variables in the data set.

\newpage

3) `track_id` should be unique to each observation. Verify this condition using `duplicated()` in R base. The argument for the `duplicated()` should be the `track_id` variable. (i) Use `dplyr::select()` to select the the `track_id` variable. (ii) Then, call `duplicated()`. (iii) Finally, call `sum()` to count the number of duplicated `track_id`. To do (i), (ii), (iii), use pipe, `%&gt;%` operator. Show the R commands.

These duplicates were included in the data set because some tracks were different in genre only.

```{r message = FALSE}
library(dplyr)
spotify %>% select(track_id) %>% duplicated() %>% sum()
```

\newpage

4) Drop the duplicates from the data set using `dplyr::filter()` and the logical argument `!duplicated()`.

```{r}
spotify <-dplyr::filter(spotify,!duplicated(track_id))
```

\newpage

5) We can explore the data set and find out that for each `track_id`, there is `genre`, `artist_name`, `track_name`, `popularity`, and some features of the track `acousticness`, `danceability`, `duration in milliseconds`, `energy`, `instrumentalness`, `key`, `liveliness`, `loudness`, `mode`,`speechiness`, `tempo`, `time_signature` and `valence`. Identify the type of each variable using `utils::str()`. We can also identify the type of each variable when we print the tibble.

```{r}
utils::str(spotify)
```

\newpage

6) Compute the average popularity of each genre. Show the R commands. How many different genres are in the data?

```{r}
library(dplyr, warn.conflicts = FALSE)
options(dplyr.summarise.inform = FALSE)
by_genre <- group_by(spotify,genre)
ave_pop <- summarise(by_genre,median= mean(popularity,na.rm = TRUE))
print(ave_pop)
```
there are 26 different genres.(there are two Children's Music, and I consider them as 1)



\newpage

7) Find the top five most popular genres based on average popularity. Choose the tracks with the top five most popular genres. Follow the steps below. Use pipe `%>%` operator.

(i) Start with the results in (6). 

(ii) Call `dplyr::arrange()` to arrange average popularity in descending order.

(iii) Call `top_n()` on the results of (ii).

(iv) Call `dplyr::select()` to select `genre`.

```{r message =FALSE}
spotify %>% select(genre,popularity) %>%
  group_by(genre) %>%
  summarise(average = mean(popularity))%>%
  dplyr::arrange(average)%>%
  top_n(5)%>%
  dplyr::select(genre)
```

\newpage
8) Create a subset of `Spotify` data by selecting the top five most popular tracks. Call the subset as `spotify_s` and print the tibble `spotify_s.`

```{r}
spotify_s <- spotify %>% dplyr::filter(genre %in% c("Dance","Hip-Hop","Rock","Rap","Pop"))
print(spotify_s)
```

\newpage

9) Use the subset `spotify_s` and explore the distribution of `genre.`
Use chunk options `out.width = "50%"`, `fig.align="center"`. Interpret the distribution of `genre` in one sentence.

```{r out.width="50%",fig.align='center'}
library(ggplot2)
ggplot(data=spotify_s) +
  geom_bar(mapping =aes(x=genre))

```
The genres of most songs are dance and hip-hop with Dance has the most songs(almost 8000) while Rap has the least songs which is less than 2000. Pop and rock have similar number of songs(around 2500).

\newpage

10) Let's explore the covariation of `genre` and `popularity`. These are mixed types of variables. Use a boxplot to explore the relationship between these two variables. It is difficult to tell that each boxplot summarizes a different number of points. Use the width of the boxplot proportional to the number of points with `varwidth = TRUE`, an argument to `geom_boxplot()`. Which genre is the most popular?

```{r}
library(ggplot2)
ggplot(data=spotify_s) +geom_boxplot(mapping = aes(x=genre, y=popularity), varwidth = TRUE)
```
Pop is the most popular genre.

\newpage
11) Boxplots were developed in an era of much smaller datasets and tend to display a prohibitively large number of `outlying values`. That you can witness from part (10). One of the remedies is using a letter value plot. Install the `lvplot` package, and try using `geom_lv()` to display the distribution of popularity vs genre. How do you interpret the plots (read the help page)?

```{r me}
library(lvplot)
ggplot(data=spotify_s)+
  geom_lv(mapping = aes(x=genre, y=popularity))

```
based on the graph above, the popularity for dance genre can be really low(the popularity could be 0 for dance genre.) the outlying popularity value for hip-hop genre and pop are similar (lower than 25 or higher than 90), however, the box below 50 is thiner for pop so we can conclude that the pop is more popular compare to hip-hop. The popularity value for rap is between 37.5 and 90. Rock is the least popular genre as the popularity value for rock is between 24 and 85.
\newpage

12) Explore the relationship between `acousticness` and `popularity`. We need to avoid overplotting for this large data set. Interpret the association between `acousticness` and `popularity.`

```{r}
library(GGally)
spotify_s%>%
  dplyr::select(acousticness,popularity)%>%
  ggpairs()
```
Since the correlation between acousticness and popularity is 0.076, we can say that there is a weak association between the two variables.

\newpage

13) Explore the association of musical features (acousticness, danceability, duration_ms, energy, instrumentalness, liveness, loudness, speechiness, tempo, valence) with popularity. We can do this either by writing ggplot() for each association or by using tidyr. Read the following [\blc blog post in R-bloggers\bc](https://www.r-bloggers.com/2016/07/plot-some-variables-against-many-others-with-tidyr-and-ggplot2/).

```{r}
library(tidyr)
library(ggplot2)
spotify_s %>% 
  select(acousticness,danceability,duration_ms,energy,instrumentalness,liveness,loudness,speechiness,popularity)%>%
  gather(-popularity,key=acousticness,value="value")%>%
  ggplot(aes(x=value,y= popularity))+geom_point()+facet_wrap(~ acousticness, scales="free")+theme_bw()
  
```

\newpage

14) Find the top five artists from the top five genres based on the most popular tracks (maximum of popularity). Explore the association between `genre` and the top five `artists.` Interpret the association. 

```{r}
spotify_s%>% arrange(desc(popularity)) %>% 
  group_by(genre) %>%
  filter(popularity == max(popularity))%>%
  filter(!duplicated(genre))%>%
  print(select(genre,artist_name,popularity))%>%
  ggplot(mapping = aes(x=genre,y=artist_name))+geom_tile(mapping=aes(fill = popularity))
```
Ariana Grande's song from dance genre has the largest popularity value of 100 while Imagine Dragons's song from Rock genre has the lowest poplarity value of 90. Daddy Yankee's song from hip-hop and Sam Smith's song from pop have similar popularity values.




